---
title: "HW3_pt4"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r include=FALSE}
rm(list = ls())
library(DAAG)                 
library(lattice)
library(MASS)
library("geneplotter")
library(ggplot2)
library(readr)
library("Hmisc")
library(corrplot)
library("dlookr")
library(dplyr)
library(tidyr)
library(ruler)
library(data.table)
library("ElemStatLearn")
library("class")
library("ISLR")
library(glmnet)
library(pls)
library(leaps)
library(tidyverse)
library(caret)
library(klaR)  
library(bestglm)
library(corrplot)
library("DataExplorer")
library('GGally')
library(boot)
```
# Q4

X and Y are commuted using by the equation given in the question using normal distribution.
```{r, echo=FALSE}
set.seed(1)
x=rnorm(100)
y=x-2*x^2+rnorm(100)
plot(x, y)
#a)
set.seed(456)

data = data.frame(x,y)
```





```{r echo = FALSE}
#Linear
linear_fit = glm(y ~ poly(x,1))
linear=cv.glm(data,linear_fit)$delta[1]
#quadratic
quadratic_fit= glm(y ~ poly(x,2))
quadratic=cv.glm(data,quadratic_fit)$delta[1]
#cubic
cubic_fit = glm(y ~ poly(x,3))
cubic=cv.glm(data,cubic_fit)$delta[1]
#quartic
quartic_fit  = glm(y ~ poly(x,4))
quartic=cv.glm(data,quartic_fit)$delta[1]

model_acc_all = c(linear,quadratic,cubic,quartic)
plot(model_acc_all, type = 'b')
```
## b)
From this plot we can see that the quadratic, cubic and quartic models are giving the almost similar and less than 0.9 . Yes this is what we expected since for the quadratic model matches the model which we used to generate the data.

The summary of the quadratic model  is shown below;
```{r echo = FALSE}
#c)
#For quadratic 
summary(quadratic_fit)
```


##C)
Statistical significance of the coefficient estimates from each model is shown below
```{r echo = FALSE}
#for linear 
confint(linear_fit)
#for quadratic 
confint(quadratic_fit)
#for cubic 
confint(cubic_fit)
#for quartic
confint(quartic_fit)
```

The confidence intervals for the coefficients of the linear and quadratic terms exclude zero ( implies that they’re significant), while the confidence intervals for the coefficients of the cubic and quartic terms(i.e. x^3 and x^4) include zero (implies that they’re not significant). The confidence interval for the intercept also excludes zero, reflecting the mean of the response.
